---
title: "Data Analysis With RMD"
format: html
editor: visual
---

I begin by doing an analysis on the various regions' real estate and rental markets. I will look at the ranges the various markets go through over a prescribed period of time. Graphical representations will be created in order to make a more in-depth analysis. I had to drill down into the data further to make better observations when some of the data appeared to be incorrect.

These are some [examples](https://github.com/Ducati013/KevinF.github.io/blob/main/Zillow%20Home%20Value%20Analysis.rmd) for a few analyses that were conducted on the [Zillow ZHVI](https://www.zillow.com/research/data/) data with R studio.

```{r echo = FALSE, message = FALSE}

library(tidyverse)
#library(stringr)
library(gridExtra)
library(grid)
library(png)
library(downloader)
library(grDevices)
library(ggplot2)

SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")

names(SingleFamilyResidenceSales) <- SingleFamilyResidenceSales %>% 
    names() %>% 
    str_replace("-05", "") #find("this","replaceWithThis")

names(SingleFamilyResidenceRental) <- gsub(pattern = "\\-..", replacement = "", x = names(SingleFamilyResidenceRental)) 
                        #get rid of the stuff after the"-" in Year 
#-------------------------------------En
```

# Wake County Home Sales

```{r echo = FALSE, message = FALSE}

#--------------------------------------Wake County Sales Eval--------------------------------------------------------------------|

WakeCountySales <- SingleFamilyResidenceSales %>% 
   filter(State %in% "NC" & CountyName %in% "Wake County") %>% #filter for all NC & Wake Co elements
  select(c(2,3,4,5,seq(from = 8, to = 277, by = 12))) %>%  # select # then, the seq from 8~277 by 12)
  pivot_longer(cols =!c(1,2,3,4), # pivot all the columns but 1:4 to yr, and val to  ....
                       names_to = 'Year',
                       values_to = ('ZHVI')) #%>% 
          #rename(City = RegionName)

WakeCountySales$Year.ordered <- factor(WakeCountySales$Year,unique(WakeCountySales$Year,level = WakeCountySales$Year))# Keeps ggplot from reorganizing points
#--------------------------------------End--------------------------------------------------------------------|

#--------------------------------------Wake Co. Point plot--------------------------------------------------------------------|

ggplot(WakeCountySales, aes(Year, ZHVI,color = RegionName,fct_reorder(x = Years)))+
  geom_point() +
  labs(title = "Wake County Home Sales")+
  scale_y_continuous(name = "Home Values (in$)", labels = scales::comma)+
  theme(strip.background = element_rect( fill = "gray"),
        plot.title= element_text(size=18, color="blue", face="bold", family = "serif"),
        axis.text.x = element_text(angle = 90,vjust = 0.5),
        axis.title = element_text(color = "blue"),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        legend.background= element_rect(fill= "gray"),
        panel.grid.major.y= element_line(color = "lightsalmon", size = 1, linetype = "dotted"),) ->p1



#---------------------------Most Expensive City Eval--------------------------------------------------------------------|

NewHillSales <- SingleFamilyResidenceSales %>% 
   filter(State %in% "NC" & CountyName %in% "Wake County" & 
            RegionName %in% c("New Hill", "Cary", "Apex","Morrisville","Rolesville","Wake Forest")) %>% #filter for all NC & Wake Co elements
  select(c(2,3,4,5,seq(from = 256, to = 277))) %>%  # select # then, the seq from 8~277 by 12)
  pivot_longer(cols =!c(1,2,3,4), # pivot all the columns but 1:4 to yr, and val to  ....
                       names_to = 'Year',
                       values_to = ('ZHVI')) #%>% 
          #rename(City = RegionName)
#--------------------------------------End--------------------------------------------------------------------|

#--------------------------Most Expensive City Point Plot-----------------------------------------------------------|
NewHillSales$Year <- factor(NewHillSales$Year,unique(NewHillSales$Year,level = NewHillSales$Year))
ggplot(NewHillSales, aes(Year, ZHVI, shape = RegionName, color = RegionName,fct_reorder(x = Years)))+ geom_point() +
  labs(title = "Most Expensive City in Wake Co.")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5),
        plot.title= element_text(size=18, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.major.y= element_line(color = "lightsalmon", size = 1, linetype = "dotted"),) +
  scale_y_continuous(name = "Home Values (in$)", labels = scales::comma) ->p2 

#--------------------Least Expensive Eval--------------------------------------------------------------------|
LeastExpensiveSales<- SingleFamilyResidenceSales %>% 
   filter(State %in% "NC" & CountyName %in% "Wake County" & 
            RegionName %in% c("Zebulon", "Wendell", "Willow Springs")) %>% #filter for all NC & Wake Co elements
  select(c(2,3,4,5,seq(from = 256, to = 277))) %>%  # select # then, the seq from 8~277 by 12)
  pivot_longer(cols =!c(1,2,3,4), # pivot all the columns but 1:4 to yr, and val to  ....
                       names_to = 'Year',
                       values_to = ('ZHVI')) #%>% 
          #rename(City = RegionName)
LeastExpensiveSales$Year <- factor(LeastExpensiveSales$Year,unique(LeastExpensiveSales$Year,level = LeastExpensiveSales$Year))
#--------------------------------------End--------------------------------------------------------------------|

#--------------------Least Expensive  Point Plot---------------------------------------------------------------|
ggplot(LeastExpensiveSales, aes(Year, ZHVI, shape = RegionName, color = RegionName,fct_reorder(x = Years)))+ geom_point()+
  labs(title = "Least Expensive City in Wake Co.")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5),
        plot.title= element_text(size=18, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.major.y= element_line(color = "lightsalmon", size = 1, linetype = "dotted"),) +
  scale_y_continuous(name = "Home Values (in$)", labels = scales::comma) -> p3
#--------------------------------------End-----------------------------------------------------------------------------------------| 

grid.arrange(p1, ncol = 1)

grid.arrange(p2, p3, ncol=1)

```

a\. What have been the overall trends in Wake County Home Values?

 - **Overall the trends have been positive.**

b\. There were dips in home values in the past 20 years. What years did these occur?

 - **Yes there were two dips in the values the first was 1997 through 1998 and the second was from 2009 through 2012.**

c\. Based on the analysis, where would be the least expensive area to purchase home? Most expensive area?

 - **Zebulon** was the **least expensive** area.

 - **New Hill** was the **most expensive** in 2018.

d\. Are any area home values trending down? Is there one area that stands out compared to others?

 - **I do not see any regions that are declining in value within Wake County.**

 - If I investigate further I do a small decrease in **New Hill** in October 2018.

 - New Hill stands out compared to the regions and it appears to be an anomaly based on historical data.

 - New Hill will need further investigation with more recent data.

 - Upon Further investigation New Hill is indeed accurate with the current data provided.

# NC Rental Market

```{r echo = FALSE, message = FALSE}

#SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
#-------------------------------------NC Regional Eval--------------------------------------------------------------------|

NcRegionRental <- SingleFamilyResidenceRental %>% 
  select(c(2,3,seq(from = 7, to = 102, by = 12))) %>%  # select # then, the seq from 7~102 by 12)      
  filter(State %in% "NC" & 
          RegionName %in% c("Asheville", "Charlotte","Durham","Fayetteville","Raleigh","Wilmington")) %>% 
  pivot_longer(cols =!c(1,2), # pivot all the columns but 1:4 to yr, and val to  ....
                       names_to = 'Year',
                       values_to = ('ZHVI')) %>% 
  arrange(ZHVI) %>% 
  rename(City = RegionName) #rename(new = old)

#-------------------------------------NC Regional Eval Plot--------------------------------------------------------------------|

#NcRegionRental$Year <- factor(NcRegionRental$Year,level = NcRegionRental$Year)
ggplot(NcRegionRental, aes(Year, ZHVI, color =City))+ geom_point()+
  labs(title = "Rental Costs in NC Cities")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5),
        plot.title= element_text(size=24, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.major.y= element_line(color = "lightsalmon", size = 1, linetype = "dashed"),
        panel.grid.minor.y= element_line(color = "sienna", size = 1, linetype = "solid")) +
  scale_y_continuous(name = "Home Values (in$)", labels = scales::comma) ->p4

ggplot(NcRegionRental, aes(Year, ZHVI,fill = City)) + geom_col(color = "black")+
  labs(title = "Rental Costs in NC Cities")+
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5),
        plot.title= element_text(size=24, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray")) +
  scale_y_continuous(name = "Home Values (in$)", labels = scales::comma) -> p5 # this adds commas to the scale labels
#-------------------------------------End--------------------------------------------------------------------|          
grid.arrange(p4, p5, ncol=1)

```

a\. What has been the overall trend in the rental market around the state? Are there any cities that have not followed this trend?

 - **The overall trend has been an increase in the rental costs in the evaluated cities.**

 - **Fayetteville** did not follow the trend, the retal rate dropped in 2011 and again in 2016 before returning to a more normal level in 2017.

b\. Where is the most expensive city to rent in? Least expensive?

 - Currently the **most expensive** city to rent in is **Asheville** while the **least expensive** is **Fayetteville.**

c\. You are trying decide between Wilmington and Asheville. Which market has the lowest rent?

 - **Wilmington** has lower rent than Asheville in 2017.

```{r, warning = FALSE, echo = FALSE, message = FALSE}
#, Asheville, Charlotte-Concord-Gastonia,Raleigh, Wilmington
#SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
#--------------------------------------NC Homes Sales--------------------------------------------------------------------|
NCHomeSales <- SingleFamilyResidenceSales %>% 
  select(c(3,4,seq(8,277,12))) %>% 
  filter(State %in% "NC" & Metro %in% c("Asheville","Charlotte-Concord-Gastonia","Raleigh", "Wilmington")) %>% 
  pivot_longer(cols = c(3:25), # select col 3 thru 25
  names_to = "Year", values_to = "ZHVI") %>% #names=colNames ,Values_to=newColName
  group_by(Metro) #by cityAreaName

#--------------------------------------NC Homes Sales Plots--------------------------------------------------------------------|
ggplot(NCHomeSales, aes(Metro, ZHVI ))+ geom_violin()+geom_boxplot(width = .1)+ #coord_flip() 
stat_summary(fun = "median", geom = "crossbar", color = "green")+
  stat_summary(fun = "mean", geom = "point", color = "orange")+
  scale_y_continuous( labels = scales::comma)+ # this adds commas to the scale labels
  labs(title = "NC Home Sales")+
  theme(plot.title= element_text(size=24, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.minor.y= element_line(color = "lightsalmon")) ->p6

ggplot(NCHomeSales, aes(Metro, ZHVI ))+geom_violin() + stat_boxplot(geom = 'errorbar') + geom_boxplot()+ coord_flip()+
  stat_summary(fun = "median", geom = "crossbar", color = "green")+
  labs(title = "NC Home Sales")+
  theme(plot.title= element_text(size=24, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.minor.x = element_line(color = "lightsalmon"),
        panel.grid.minor.y = element_line(color = "cornflowerblue")) ->p7# it will not give the the minors
#--------------------------------------end-----------------------------------------------|
grid.arrange(p6, p7, ncol=1)
```

a\. According to the results, which market has the lowest median price (represented as horizontal bar in

box plot)?

 - **Charlotte-Concord-Gastonia** had the lowest median price.

b\. The violin plot will show density meaning the wider the plot is, the more observations occur within

that area. Which market has the most density around the median value of homes?

 - **Charlotte-Concord-Gastonia** once again has the largest density around the median value in that metro area.

# Relocation Home Value Comparison

```{r ,warning = FALSE, message =FALSE, echo= FALSE}

#--------------------------------------National Home Sales Eval-------------------------------------------------------------------|

NationalHomeSales <- SingleFamilyResidenceSales %>% 
  select(c(2,3,4,seq(8,277,12))) %>% 
  filter(RegionName %in% c("Chicago","Denver","Houston","New York") & State %in% c("IL","CO","TX","NY")) %>% 
  pivot_longer(cols = c(4:26), names_to = "Year", values_to = "ZHVI")


ggplot(NationalHomeSales, aes(Year, ZHVI,color = RegionName))+
  geom_point()+
  facet_wrap(~Metro)+  # this creates 4 plots,one for each city to compare :)
  labs(title = "City Home Value Comparison")+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust=0.5, size=5),
        plot.title= element_text(size=18, color="blue", face="bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", colour="black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.major.y= element_line(color = "lightsalmon"),)+
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)+
  geom_vline(xintercept = c(17,23), linetype = "dashed")->p8
#--------------------------------------end-----------------------------------------------|

#--------------------------------------more in depth eval for past 5 years-----------------------------------------------
Past5YearsNationalHomeSales <- SingleFamilyResidenceSales %>% 
  select(c(2,3,4,seq(184,277,6))) %>% 
  filter(RegionName %in% c("Chicago","Denver","Houston","New York") & State %in% c("IL","CO","TX","NY")) %>% 
  pivot_longer(cols = c(4:19), names_to = "Year", values_to = "ZHVI") %>% 
  rename(City = RegionName)


ggplot(Past5YearsNationalHomeSales, aes(Year, ZHVI,shape = City,color = City))+
  geom_point()+
  #facet_wrap(~Metro)+  # this creates 4 plots,one for each city to compare :)
  labs(title = "City Home Value Comparison")+
  theme(axis.text.x = element_text(family = "mono", colour = "black", face = "bold", angle = 90, vjust=0.5, size=8),
        plot.title = element_text(size = 18, color = "blue", face = "bold", family = "serif"),
        axis.title = element_text(color = "blue", size = 12),
        axis.text = element_text(family = "mono", color = "black", face = "bold"),
        plot.background = element_rect(fill = "gray"),
        panel.grid.major.y= element_line(color = "lightsalmon"),)+
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)+
  geom_vline(xintercept = c(3,5,16), linetype = "dashed")->p9

  



Chicago= (208500-151700)/(2018-2013) #(y2 - y1)/(x2 - x1) = m
Houston= (185800-123200) /(2018-2013)
# Chicago<Houston
Change <- if(Chicago < Houston){ 
  print("Chicago") 
} else {
  print("Houston")
}

grid.arrange(p8, p9, ncol=1)
```

a. Based on your analysis, which city's housing is most affordable? Least affordable?

 - **Houston-The Woodlands-Sugar Land** appears to be the most affordable, and **New York-Neward-Jersey City** are the least affordable.

b. Which cities saw the largest change in prices over the past 5 years? Which city has remained more

consistent (i.e., no huge swings up or down in home values)?

 - During the past 5 years **New York-Neward-Jersey City**, and **Denver-Aurora-Lakewood** saw the largest change in prices.

 - **Houston-The Woodlands-Sugar Land** has the most consistent housing market since 1996, and **`r Change`** has had the most consistent housing market during over the past 5 years.

c. During the market downturn in 2012, which cities were most impacted? Which cities have

recovered?

 - **New York-Newark-Jersey City** and **Chicago-Naperville-Elgin** were the most effected by the 2012 downturn.

 - All the cities have recovered since the 2012 market downturn.
