---
title: "Pivoting Tables"
format: html
editor: visual
---

```{r, message =FALSE, echo = FALSE}

library(tidyverse)
library(gridExtra)
library(grid)
library(png)
library(downloader)
library(grDevices)
library(ggplot2)

tuition_cost <- read_csv("C:/Users/ducat/Documents/UNCW_Grad_school/MIS 503 R for DataScience_Programming/module5/Mod5Assignment2/tuition_cost.csv")
#tuition_cost
```

## Creating tidy data

-   This is an exercise showcasing some of the skills I have learned thus far.

I will begin with reading the csv file into R and perform some information gathering from the data. I begin the investigation by finding there are significant number of columns that are date based. I will merge those date based columns into one column and their corresponding values and merge them into one column thus converting the data into a more useful format. The new name for the years column will be 'Year' and the corresponding values will be 'tuition'. I will be using piping to make the code tidier.

```{r, message = FALSE, echo = TRUE}

#, merge 4 col to 2, separate the tuition col to InState and OutState, change year to int

Tuition <- pivot_longer(tuition_cost                         # this will take 4 columns and merge into 1
                        , c('2017','2018','2019','2020'),    # these are the 4 columns chosen
                        names_to = 'Year',                   # this is the new name for the columns
                        values_to = 'tuition') %>%           # and their values will be assigned to this column
           separate(tuition,                                 # then we are going to separate the elements in 'tuition' col
              into=c("InState","OutState"),                  # those elements will be separated by '/' into 'InState' and 'OutState'
              sep='/',convert=TRUE) 
#Tuition              
Tuition$Year <- as.integer(Tuition$Year)

```

## North Carolina Schools

This section will filter down all the observations to a North Carolina School with a 4 year degree path calculating tuition rate as in-state for this evaluation. As as secondary evaluation all the schools with University of North Carolina in the name will be evaluated. All of these evaluations will be plotted with the year on the x-axis and the tuition rate on the y-axis. I will add some titles to the evaluations and color the universities to make the graphs easier to evaluate.

```{r echo= FALSE, message = FALSE}

# , filtering down the Tuition to Nc instate 4 year school and plot it, then a unc sys and plot

Public <- Tuition %>% 
  filter(state_code == "NC" & type == "Public" & degree_length == "4 Year" )   #filter for NC,Public,4 year schools

ggplot(Public, aes( Year, InState,color = name))+
  geom_point()+
  labs(title = "In-StateTuition for North Carolina Schools", x = "Year", y = "Cost in $")->p1

#ggplot(Public, aes(Year,InState,fill = name))+    # just to see if I can
 # geom_col()

UNC <- filter(Public,grepl("University of North Carolina",name))  # grpl searchs for the "" in the "name" col in DF Public

ggplot(UNC, aes(Year,InState, color = name))+
  geom_point()+
  labs(title = "In-State Tuition for UNC Schools", x = "Year", y = "Cost in $")+
  theme(legend.position = "bottom", legend.title = element_blank())+   # legend on the bottom
  guides(color = guide_legend(nrow = 4))->p2   # }


grid.arrange(p1, ncol = 1)
grid.arrange(p2, ncol = 1)

```

## Additional Universities located in South Carolina and Colorado:

The next section will begin with an evaluation for a public Colorado school with a 4 year degree path at an in-state rate for calculations. The next evaluation will be for a public South Carolina school with a 2 year program at an out-of-state tuition rate. I will add a title to the graphs with 'year' for the x-axis and 'tuition-rates' for the y-axis for both graphs.

```{r ,echo= FALSE, message= FALSE}

CO <- Tuition %>% 
  select(everything(),-InState) %>% 
  filter(state_code == "CO" & type == "Public" & degree_length == "4 Year")
  
SC <- Tuition %>% 
  select(everything(),-OutState) %>% 
  filter(state_code == "SC" & type == "Public" & degree_length == "2 Year")
  

ggplot(CO, aes(Year, OutState, color = name))+
  geom_point()+
  labs(title ="Colorado 4Year Public School, OutState Tuition", x= "Year", y = "Cost in $")+
  theme(legend.position = "right", 
        legend.title = element_blank() # +  #legends on the bottom with not legend title
  # guides(color = guide_legend(nrow = 4))   # This makes 4 rows of Universities
        )->p3

ggplot(SC,aes(Year,InState, color = name))+
  geom_point()+
  labs(title = "S.Carolina 2Year Public School, InState Tuition", x = "Year", y = "Cost in $")+
  theme(legend.position = "right", 
        legend.title = element_blank(),
        legend.text = element_text(size = 8)
        #legend.text = element_text(angle = 0,vjust=0, hjust=0)  # this will move the legend text up, down, and angle it
        )->p4

grid.arrange(p3, ncol = 1)
grid.arrange(p4, ncol = 1)

```
